---
title: "simpleRegress"
author: "Gianna Chevez"
date: "Fall 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 



## Model WEIGHT ~ HEADLEN  this is an example  

```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```

Best to worst 

1. Neck          ARsard: 0.87        RST: 43.91
2. Headlen       ARsard: 0.69        RST: 67.81

The Neck model has a higher ARsard (0.87) and a lower RST (43.91), indicating better performance and lower error. The Head length model has a lower ARsard (0.69) and a higher RST (67.81), showing weaker performance and greater error.


## Model WEIGHT ~ HEADWTH  this is an example  

```{r}
headwthmodel <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(headwthmodel)
summary.lm(headwthmodel)
```

Best to worst 

1. Neck                ARsard: 0.87        RST: 43.9

2. Head length         ARsard: 0.69        RST: 67.81

3. Head width          ARsard: 0.60        RST: 76.41

The Neck model ranks highest with the best performance: ARsard of 0.87 and the lowest RST of 43.91. The Head length model is in the middle, with a lower ARsard of 0.69 and a higher RST of 67.81. The Head width model ranks worst, having the lowest ARsard of 0.60 and the highest RST of 76.41, indicating the weakest performance and highest error.


## Model WEIGHT ~ LENGTH  this is an example 

```{r}
lengthmodel <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(lengthmodel)
summary.lm(lengthmodel)
```

Best to worst 

1. Neck                ARsard: 0.87        RST: 43.9

2. Length              ARsqrd: 0.74        RST: 61.83

3. Head length         ARsard: 0.69        RST: 67.81

4. Head width          ARsard: 0.60        RST: 76.41

The Neck model remains the best performer with the highest ARsard (0.87) and the lowest RST (43.91). Length is second, with a solid ARsard of 0.74 and moderate RST of 61.83. Head length ranks third, with a lower ARsard (0.69) and a higher RST (67.81). Head width ranks last, with the lowest ARsard (0.60) and the highest RST (76.41).

## Model WEIGHT ~ AGE  this is an example 

```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```

Best to worst 

1. Neck                ARsard: 0.87        RST: 43.9

2. Length              ARsqrd: 0.74        RST: 61.83

3. Head length         ARsard: 0.69        RST: 67.81

4. Head width          ARsard: 0.60        RST: 76.41

5. Age                 ARsard: 0.55        RST: 81.47

Neck is still the best-performing model with the highest ARsard (0.87) and the lowest RST (43.91). Length comes in second, with a decent ARsard (0.74) and moderate RST (61.83). Head length ranks third with a lower ARsard (0.69) and a higher RST (67.81). Head width ranks fourth, with a further drop in ARsard (0.60) and an RST of 76.41. Age ranks worst, with the lowest ARsard (0.55) and the highest RST (81.47), indicating the weakest performance and largest error.
